---
title: "R Coding Exercise"
---

Mojisola Popoola

```{r}
library("dplyr")
library("tidyverse")
library("ggplot2")
library("dslabs")
library("renv")
```

```{r}
#look at help file for gapminder data
help(gapminder)
#get an overview of data structure
str(gapminder)
```

```{r}
#get a summary of data
summary(gapminder)
```

```{r}
#determine the type of object gapminder is
class(gapminder)
```

```{r}
africadata = filter(gapminder, continent == 'Africa')
head(africadata)
```

# To get overview of africadata

```{r}
str(africadata)
```

#Summary of african data

```{r}
summary(africadata)
```

# Infant survival rate

```{r}
infant_survival_rate = africadata %>%
  select(infant_mortality, life_expectancy)
head(infant_survival_rate)
```

# Structure of infant survival

```{r}
str(infant_survival_rate)
```

# summary of infant survival

```{r}
summary(infant_survival_rate)
```

# Infant mortality

```{r}
african_mortality = africadata %>%
  select(population, life_expectancy)
head(african_mortality)
```

#Part 1 - Amelia Tristan

```{r}
#installed and calling dslabs, tidyverse
library(dslabs)
library(tidyverse)
library(ggplot2)

#open help page for gapminder data and explore the data structure, class, and contents
help(gapminder) #help page
str(gapminder) #structure
summary(gapminder) #contents
class(gapminder) #class

gapminder
```

```{r}
#filtering gapminder data to include only Africa
africadata = dplyr::filter(gapminder, continent == "Africa")
str(africadata) #explore structure
summary(africadata) #explore filtered data
```

```{r}
#selection of variables from africadata
ad1 = dplyr::select(africadata, infant_mortality, life_expectancy) #selecting infant_mortality and #life expectancy

ad2 = dplyr::select(africadata, population, life_expectancy) #selecting pop and life expectancy

#exploring structure and data contents for each new data set
str(ad1) #structure for ad1
str(ad2) #structure for ad2
summary(ad1) #content summary for ad1
summary(ad2) #content summary for ad2
```

```{r}
plot(ad1$infant_mortality, ad1$life_expectancy)
#plot(ad2$population, ad2$life_expectancy)

plot2 = ggplot(gapminder, aes(log10(population), life_expectancy)) + geom_point()
plot2
```

As infant mortality rates increase, life expectancy decreases. As population increases, life expectancy increases, but only up to a certain population amount before the population becomes detrimental to life expectancy and "resets."

```{r}
#filter and select null infant mortality years in africa data
adna <- africadata %>% dplyr::filter(is.na(infant_mortality)) %>% dplyr::select(year)
adna$year #print all years with null infant mortality

#filtering to only the year 2000
ad2000 <- africadata %>% dplyr::filter(year == 2000)
summary(ad2000) #summary of contents
str(ad2000) #structure
```

```{r}
#plotting infant mortality by life expectancy for year 2000
plot(ad2000$infant_mortality, ad2000$life_expectancy)
#plotting life population by life expectancy for year 2000 on population axis scaled by log10
plot3 = ggplot(ad2000, aes(log10(population), life_expectancy)) + geom_point()
plot3 #print plot
```

```{r}
#life_expectancy~infant_mortality model fit for year 2000
inf_le_fit = lm(life_expectancy~infant_mortality, data = ad2000)
##life_expectancy~population model fit for year 2000
pop_le_fit = lm(life_expectancy~population, data = ad2000)

summary(inf_le_fit)
summary(pop_le_fit)
```

## The R-squared is 0.47 which means that 47% variation in life expectancy can be explained by infant mortality

I created two linear models using lm() function with life expectancy as the dependent variable for both infant mortality and population as the single predictor. Based on the fitted models, population is more significant than infant mortality, with a corresponding high r-squared value. The best model would be the population model.

```{r}
regression = lm(african_mortality$life_expectancy ~ african_mortality$population)
rsq <- function(x, y) summary(lm(african_mortality$life_expectancy ~ african_mortality$population))$r.squared
r_square = round(rsq(obs, mod), 3)

coeff = round(summary(regression)$coefficients[2],9)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  african_mortality$population,
  african_mortality$life_expectancy,
  main='African Life Expectancy \n Population Effect',
  xlab='Population',
  ylab='Life Expectancy',
  col='pink', 
  bg='purple', 
  pch = 21,
)
```

# Missing years for infant mortality

```{r}
missing_infant_mort_rows = africadata %>%
  filter(is.na(infant_mortality))
head(missing_infant_mort_rows)
```

# Perform year group by to identify missing year

```{r}
missing_infant_mort_yrs = missing_infant_mort_rows %>%
  group_by(year) %>%
  count(infant_mortality)
missing_infant_mort_yrs
```

# Extract only year 2000 from data

```{r}
africadata2000 = filter(africadata, year == 2000)
africadata2000
```

# Structure of year 2000 data

```{r}
str(africadata2000)
```

#Part 2 - Mojisola Popoola \# Summary of year 2000 data

```{r}
summary(africadata2000)
```

# To create a filter from dataset

```{r}
african_mortality2000 = africadata2000 %>%
  select(population, life_expectancy)
head(african_mortality2000)
```

# To create another dataset

```{r}
infant_survival_rate2000 = africadata2000 %>%
  select(infant_mortality, life_expectancy)
head(african_mortality2000)
```

# infant mortality effect

```{r}
regression = lm(infant_survival_rate2000$life_expectancy ~ infant_survival_rate2000$infant_mortality)
rsq <- function(x, y) summary(lm(infant_survival_rate2000$life_expectancy ~ infant_survival_rate2000$infant_mortality))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],3)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  infant_survival_rate2000$infant_mortality,
  infant_survival_rate2000$life_expectancy,
  main='African Life Expectancy \n Infant Mortality Effect',
  xlab='Infant Mortality',
  ylab='Life Expectancy',
   col='green', 
   bg='violet', 
   pch = 21,
)
```

As the infant mortality increases the life expectancy decreases

# Population effect

```{r}
regression = lm(african_mortality2000$life_expectancy ~ african_mortality2000$population)
rsq <- function(x, y) summary(lm(african_mortality2000$life_expectancy ~ african_mortality2000$population))$r.squared
r_square = round(rsq(obs, mod), 3)
# summary(regression)$coefficients - keep to sanity check
coeff = round(summary(regression)$coefficients[2],9)
y_int = round(summary(regression)$coefficients[1],3)
plot(
  african_mortality2000$population,
  african_mortality2000$life_expectancy,
  main='African Life Expectancy \n Population Effect',
  xlab='Population',
  ylab='Life Expectancy',
  col='violet', 
  bg='red', 
  pch = 21,
)
```

# Simple model fit

# Life expectancy to infant mortality

```{r}
fit1 = lm(life_expectancy ~ log(infant_mortality), data=infant_survival_rate2000)
summary(fit1)
```

The R-squared is 0.5411, this means that 54% of life expectancy can be explained by infant mortality

# Fit life expectancy to population

```{r}
fit2 = lm(life_expectancy ~ log(population), data=african_mortality2000)
summary(fit2)
```

The R-squared is 0.01044 which means that 1% of population can be explained by life expentancy

Part2 with another dataset- Mojisola Popoola

```{r}
library(ggplot2)
library(tidyverse)
```

```{r}
help(txhousing)
head(txhousing)
```

```{r}
summary(txhousing)
```

≈

```{r}
```

```{r}
str(txhousing)
```

```{r}
year_txhousing = filter(txhousing, year == '2000')
head(year_txhousing)
```

```{r}
summary(year_txhousing)
```
